DATA 621 Week #3 Textbook Exercises

### 2019-10-05 DATA621/Week 03 Textbook Exercises.Rmd
Worked on DATA 621 HW 3

---
title: "DATA 621 Week #3 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(ggthemes)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

@@

@@

>>>>

---
title: "DATA 621 Week #3 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
---

```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(ggthemes)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

@@

For the `fat` data used in this chapter, a smaller model using only `age`, and `weight`, `height` and `abdom` was proposed on the grounds that these predictors are either known by the individual or easily measured.

a. Compare this model to the full thirteen predictor model used earlier in the chapter.  Is it justifiable to use the smaller model?

```{r}
fat <- read.delim('./data/fat.txt') 
full <- lm(brozek ~ age + weight + height + neck + chest + abdom + hip + thigh + knee + ankle + biceps + forearm + wrist, data = fat)
small <- lm(brozek ~ age + weight + height + abdom, data = fat)
summary(small)
```

b. Compute a 95% prediction interval for median predictor values and compare to the results to the interval for the full model.  Do the intervals differ by a practically important amount? **No it doesn't differ.  See below:**

### Small Model

```{r}
x0 <- data.frame(t(apply(model.matrix(full), 2, median)))
predict(small, x0, interval = "prediction")
```

### Full Model

```{r}
predict(full, x0, interval = "prediction")
```

c. For the smaller model, examine all the observations from case number 25 to 50.  Which two observations are particularly anomalous? **Case 25 and 33.**

```{r}
data.frame(actual = fat[25:50, "brozek"], yhat = round(predict(small, fat)[25:50], 1)) %>% 
  rownames_to_column() %>%
  mutate(error = abs((yhat - actual)/yhat)) %>% 
  arrange(desc(error)) %>%
  top_n(5) %>%
  kable() %>% 
  kable_styling()
```

d. Recompute the 95% prediction interval for the median predictors after these two anomoulous cases have been excluded from the data.  Did this make much difference to the outcome? **It certainly did.  See below:**

```{r}
fat <- fat[!row.names(fat) %in% c(25,33),]

full <- lm(brozek ~ age + weight + height + neck + chest + abdom + hip + thigh + knee + ankle + biceps + forearm + wrist, data = fat)
small <- lm(brozek ~ age + weight + height + abdom, data = fat)

x0 <- data.frame(t(apply(model.matrix(full), 2, median)))
```

### Small Model

```{r}
predict(small, x0, interval = "prediction")
```

### Full Model

```{r}
predict(full, x0, interval = "prediction")
```

@@

Use the `odor` dataset with odor as the response variable and `temp` as the predictor. Consider all possible models that also include all, some or none of the other two predictors.  Report the coefficient for temperature, its standard error, t-statistic and p-value in each case. Discuss what stays the same, what changes and why.  Which model is the best?

```{r}
odor <- read.delim("data/odor.txt")
models <- c("odor ~ temp", "ordor ~ temp + gas", "ordor ~ temp + pack", "ordor ~ temp + gas + pack")
coefficients <- matrix(NA, 4, 6)

fit <- lm(odor ~ temp, data = odor)
coefficients[1,] <- c("odor ~ temp", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
fit <- lm(odor ~ temp + gas, data = odor)
coefficients[2,] <- c("odor ~ temp + gas", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
fit <- lm(odor ~ temp + pack, data = odor)
coefficients[3,] <- c("odor ~ temp + pack", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
fit <- lm(odor ~ temp + gas + pack, data = odor)
coefficients[4,] <- c("odor ~ temp + gas + pack", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
coefficients <- as.data.frame(coefficients)
names(coefficients) <- c("Formula", names(summary(fit)$coefficients[2,]), "Adjusted R-squared")
coefficients %>%
  kable() %>%
  kable_styling()
```
